---
title: "Tidying and Transforming Data"
author: "<p><img class='dj'  data-pin-nopin='true'> Dhananjay Kumar</p>"
date: "September 28, 2016"
output:
  html_document:
   theme: flatly
   highlight: pygments
   toc: true
   toc_float: true
   css: dj.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(plotly)
library(knitr)
library(magrittr)
library(sqldf)
library(stringr)
library(tcltk)
library(DT)
```

## Assignment

**1.** Create a .CSV file that includes the Airline information .

**2.** Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

**3.** Perform analysis to compare the arrival delays for the two airlines.

## Load Data

### 1.1 Load CSV file
Load CSV file from desktop and validate it using Head.
```{r warning=FALSE}
untidyData <-  read.csv(paste0("C:/data/tidyingData.csv"), header=T)
kable(untidyData)
```



## Tidy Data 

### 2.1 Tidy Data - I

Lets tidy the untidy dataset by gathering the destination cities in one Column : ** Destination**.

```{r warning=FALSE}
untidyData <- untidyData %>% gather(Destination, n, Los.Angeles:Seattle)
datatable(untidyData)
```

### 2.2 Tidy Data - II

Lets create column names from Categorical data **Status**. This will make our untidy dataset to Tidy dataset as all varibales will be moved to Columns and  Observations into Rows.

```{r warning=FALSE}
tidyData <- untidyData %>% spread(Status, n)
datatable(tidyData)
```


## Analysis

### 3.1 Probability

Lets add two more columns to our tidy dataset : **onTime_Probability**, **Delayed_Probability**
```{r warning=FALSE}
tidyData$onTime_Probability <- round((tidyData$onTime / (tidyData$delayed + tidyData$onTime)), digits = 3)
tidyData$Delayed_Probability <- round((tidyData$delayed / (tidyData$delayed + tidyData$onTime)), digits = 3)
datatable(tidyData)
```

### 3.2 Summarize Data

Below we have summarized data on the basis of On time Probability.
```{r warning=FALSE}
b <- data.frame((summary(sqldf('select onTime_Probability from tidyData where AirLine = "ALASKA"'))), summary(sqldf('select onTime_Probability from tidyData where AirLine = "AM WEST"')) )
AL <- str_split_fixed(b$Freq, ":", 2)
AMWest <- str_split_fixed(b$Freq.1, ":", 2)
meanData <- data.frame(AL, AMWest)
meanData$X1.1 <- NULL
colnames(meanData) <- c("Function", "Alaska", "AM West")
kable(meanData)

```


As seen above, the Alaskan Airlines has better on Time performance than AM West.

## Plots {.tabset}

### 4.1 Scatter Chart

The Scatter Chart below shows that Phoenix has best on time probability for both AirLines and San Francisco has least on time probability for both Airlines.
```{r}
bsc <- ggplot(tidyData, aes(x = Destination , y = onTime_Probability))  + geom_point(aes(color = onTime_Probability, size = onTime_Probability, shape = factor(AirLine))) +  scale_colour_gradient(low = "purple")
ggplotly(bsc)
```


### 4.2 Density Chart

The Density Chart below support our analyses. The Alaskan Airlines has better Ontime performance for almost all the destinations.
```{r}
dPlot <- qplot(onTime_Probability, data=tidyData, geom='density', color=AirLine, xlim =c(0.50, 1)) + facet_grid(Destination ~.)
ggplotly(dPlot)
```



