---
title: "hw3_dkumar"
author: "Dhananjay Kumar"
date: "February 09, 2017"
output: html_document
---

##  Problem set 2

Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.
Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer. Please submit your response in an R Markdown document using our class naming
convention, E.g. LFulton_Assignment2_PS2.png

### Factorization function

```{r cars}
myfactorize  <- function(A) {
  
  #  Array size
  d =dim(A)[1]
  d
  
  # create an identity matrix
  y <- diag( dim(A)[1])
  y
  
  # initializing parameters 
  
  temp = y
  temp[2,1] <- -A[2,1]/A[1,1]
  ltemp= temp
  
  # initialize the Upper elimination matrix 
  temp= temp %*% A
  L=ltemp
  i=1
  
  # initialize the lower elimination matrix
  L= solve(L)

   # looping through the martix columns  
  for  (j in 1:(d-i)) {
    #y <- diag( dim(A)[1])
    for  (i in 3:d) {
      # check if the pivot element. no permutaion is done when pivot is 0. 
      if (i != j) {
        # Check if the element is zero
        if (temp[i,j] != 0 ) {  
          if (temp[j,j] < 0)  y[i,j] <- -temp[i,j] / temp[j,j]     
          if (temp[j,j] > 0  )  y[i,j] <- -temp[i,j]/ temp[j,j]
          
          #place holder for the Uppper matrix U
          temp= y%*%temp
          
          # place holder for the Lower matrix L
          L= L %*%solve(y)
          
          #print(temp). Please uncomment to see all the Upper elimination matrices
          #print(L). Please uncomment to see all the Lower elimination matrices

          #reset y to the identity matrix
          y <- diag( dim(A)[1])
          
        }
        
      }
      
    }
  } 
  print("Upper Matrix U")
  print(temp)
  print("Lower Matrix L")
  print(L)
  U=temp
  solution <- L%*%U
  print("Compare LU with original Matrix A. Below is the LU matrix:  ")
  print (solution) 
  print("Below is the original Matrix A: ")
  print(A)
  return(all.equal(solution,A))  
  
}
```


### Call function

```{r}
A <- matrix(c(1,2,3,1,1,1,2,0,1),nrow=3)
myfactorize(A)
```

