---
title: "Tidying and Transforming Data"
author: "<p><img class='dj'  data-pin-nopin='true'> Dhananjay Kumar</p>"
date: "September 28, 2016"
output:
  html_document:
   theme: flatly
   highlight: pygments
   toc: true
   toc_float: true
   css: dj.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(plotly)
library(knitr)
library(magrittr)
library(sqldf)
library(stringr)
library(tcltk)
library(DT)
```

## Assignment

**1.** Create a .CSV file that includes the Airline information .

**2.** Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

**3.** Perform analysis to compare the arrival delays for the two airlines.

## Load Data

### 1.1 Load CSV file
Load CSV file from desktop and validate it using Head.
```{r warning=FALSE}
untidyData <-  read.csv(paste0("C:/data/China.csv"), header=T)
datatable(untidyData)
```



## Tidy Data 

### 2.1 Tidy Data - I

Lets tidy the untidy dataset by gathering the Years in one Column : **Year**.

```{r warning=FALSE}
untidyData <- untidyData %>% gather(Year, Value, X2000:X2014)
untidyData$Series.Code <- NULL
head(untidyData)
```

### 2.2 Remove unwanted Columns

Since this data is for China, there is no need to mention that in an explicit Column.

```{r}
untidyData$Country.Name <- NULL
untidyData$Country.Code <- NULL
```

### 2.3 Tidy Data - II

Lets create column names from Categorical data **Series.Name**. This will make our untidy dataset to Tidy dataset as all varibales will be moved to Columns and  Observations into Rows.

```{r warning=FALSE}
tidyData <- untidyData %>% spread(Series.Name, Value)
datatable(tidyData)
```

### 2.3 Summarize

```{r warning=FALSE}
chinaGDP <- data.frame(tidyData$Year, tidyData$`GDP (current US$)`, tidyData$`GDP growth (annual %)`, tidyData$`Agriculture, value added (% of GDP)`)
mutate(chinaGDP, Agriculture_Current = tidyData$`GDP (current US$)`*(tidyData$`Agriculture, value added (% of GDP)`/100))
colnames(chinaGDP) <- c("GDP-Current", "GDP-Growth","Agriculture as % of GDP")

datatable(chinaGDP)
```


