---
title: "Web APIs"
author: "Dhananjay Kumar"
date: "October 30, 2016"
output:
  html_document:
   theme: flatly
   highlight: pygments
   css: dj.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(jsonlite)
library(dplyr)
library(knitr)
```

## Assignment Summary

Our task  for the week 9 assignment is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and
transform it to an R dataframe.

For this assignment I opted for the Best Seller Books API from New York Times.

## Load URL

Lets load the NYT API and parse it R. For parsing json api, I am using **JSONLITE** Library.

```{r warning=FALSE}
json_file <- "https://api.nytimes.com/svc/books/v3/lists/best-sellers/history.json?api-key=5fe82da6fd5740a7a90d749aa086a9b5"
json_data <- fromJSON(paste(readLines(json_file), collapse=""))
```

## Create Data frame 

Create Data frame for the data retreived in the above step. Please not that I am using **list()** function of DPLYR package to append data when parsed through FOR Loop.
```{r warning=FALSE}
listTab <- list()
rankTab <- list()
dataBestSeller <- length(json_data$results)
rankData <- length(json_data$results)
for (i in 1:length(json_data$results)) {
  rankData$Title <- dataBestSeller$Title <- gsub( "#", "", json_data$results$title[i])
  dataBestSeller$Description <- json_data$results$description[i]
  dataBestSeller$Contributor <- json_data$results$contributor[i]
  dataBestSeller$Author <- json_data$results$author[i]
  dataBestSeller$Publisher <- json_data$results$publisher[i]
  dataBestSeller$isbn10 <- json_data$results$isbns[[i]]$isbn10
  
#  rankData$Title <- gsub( "#", "", json_data$results$title[i])
  rankData$Rank <- json_data$results$ranks_history[[i]]$rank
  rankData$Weeks_On_List <- json_data$results$ranks_history[[i]]$weeks_on_list
  rankData$ISBN10 <- json_data$results$ranks_history[[i]]$primary_isbn10
  rankData$ListName <- json_data$results$ranks_history[[i]]$list_name
  rankData$Publication_Date <- json_data$results$ranks_history[[i]]$published_date
  rankData$Best_Seller_Date <- json_data$results$ranks_history[[i]]$bestsellers_date
  
  dfBooks <- data.frame(dataBestSeller)
  rankBooks <- data.frame(rankData)
  
  listTab[[i]] <- dfBooks
  rankTab[[i]] <- rankBooks
}
 opBooks <- bind_rows(listTab)
 opBooks$X11L <- NULL
 
 opRank <- bind_rows(rankTab)
 opRank$X11L <- NULL
 
 kable(opBooks)
# kable(opRank)
```

